Precise Data Flow Analysis\\in the Presence of Correlated Method Calls

### 2015-03-08 sas2015/content/example.tex
minor changes

In summary, the example discussed in this section shows how the precision of 
IFDS-based taint analysis can be improved by taking into account how paths
become infeasible in the presence of correlated calls. In particular, we have
shown how a false positive reported by a standard IFDS-based algorithm is eliminated.
While the discussion in this section has focused on the specific problem of taint analysis,
we would like to emphasize that our technique generally applies to \textit{any}
data-flow-analysis problem that can be expressed in the IFDS framework. This includes
many common analysis tasks such as reaching definitions, constant propagation, slicing,
typestate analysis, pointer analysis, and lightweight
shape analysis.
%\ftodo{Ondrej, can you give a few more/better examples?}

\subsection{Occurrences of Correlated Calls}
How often do correlated calls occur in practice? To assess the benefit of the correlated-calls analysis, we counted the number of correlated calls that occur in programs of the Dacapo benchmarks~\cite{blackburn2006dacapo}, using the WALA framework for static analysis on Java bytecode~\cite{fink2012wala}.
Our goal was to obtain an upper bound on the number of redundant IFDS-result nodes that could be potentially removed by our analysis. The results are shown in Table~\ref{tab:dacapostat} in the Appendix.

@@

\subsection{An Example from the Scala Collections Library}\label{sec:collex}
The Scala collections library contains the trait \code{TraversableOnce} that is
shared by both collections and iterators over them. 
%\ftodo{
%  I think it would be helpful to show a little more of the code of the trait here,
%  such as the trait declaration and the signature of the toArray method containing the
%  correlated calls. You can abbreviate everything else with ...
%}
The \code{toArray} method of this trait
creates an array and copies the contents of the collection or iterator into it:
\inputMintedNoLNs{scala}{collectionsExample.scala}
%\otodo{Marianna, can you figure out how to get LaTeX to syntax-highlight this?} 
When \code{this} refers to an iterator rather than a collection, the call to \code{this.size}
extracts all elements of the iterator to count them.
%\ftodo{
%  Would it make sense to show the code for the size method, and refer to the line numbers
%  of that method in the above sentence?  
%}
 At the call to \code{copyToArray},
the iterator is already empty, so nothing is copied to the newly created array.
%\ftodo{
%Would it make sense to show the code for copyToArray here?
%}
One could design an IFDS analysis to detect this kind of bug.

However, the implementation of \code{TraversableOnce.toArray} is actually correct because the above
code is guarded with a test:
\inputMintedNoLNs{scala}{isTraversable.scala}
%\ftodo{It is confusing that this code fragment is labeled with line number 19 although it presumably
%  occurs in the code before the previously discussed two lines. I think it is better to show
%  all of the code at once, in a figure, and include at least the headers of all the methods mentioned
%  here.}
When the \code{isTraversableAgain} method returns false, as it does for
an iterator, the \code{toArray} method uses a different
(less efficient) implementation. The bug report would therefore be a false positive. 
The \code{isTraversableAgain} method is easy to analyze: it returns the constant
true in a collection and the constant false in an iterator. However, in order
to eliminate the false positive bug report, an analysis would need to rule out infeasible
paths using correlated calls. Specifically, the following path triggers the bug, but
is infeasible: first, call \code{isTraversableAgain} on a collection, returning
true, then call \code{size} and \code{copyToArray} on an iterator. Our correlated
calls analysis could determine that this path is infeasible because it calls the
collection version of \code{isTraversableAgain} but the iterator versions of
\code{size} and \code{copyToArray}.
%\ftodo{Would it be helpful to create a diagram of the relevant fragments of the supergraph,
%with the infeasible path in bold?}

\commentout{
    For reference, here is the full code of the method:

def toArray[B >: A : ClassTag]: Array[B] = {
    if (isTraversableAgain) {
        val result = new Array[B](size)
        copyToArray(result, 0)
        result
    }
    else toBuffer.toArray
}

}

>>>>

In summary, the example discussed in this section shows how the precision of 
IFDS-based taint analysis can be improved by taking into account how paths
become infeasible in the presence of correlated calls. In particular, we have
shown how a false positive reported by a standard IFDS-based algorithm is eliminated.
While the discussion in this section has focused on the specific problem of taint analysis,
we would like to emphasize that our technique generally applies to \textit{any}
data-flow-analysis problem that can be expressed in the IFDS framework. This includes
many common analysis tasks such as reaching definitions, constant propagation, slicing,
typestate analysis, pointer analysis, and lightweight
shape analysis.

\subsection{Occurrences of Correlated Calls}\vspace{-.5mm}
How often do correlated calls occur in practice? To assess the benefit of the correlated-calls analysis, we counted the number of correlated calls that occur in programs of the Dacapo benchmarks~\cite{blackburn2006dacapo}, using the WALA framework for static analysis on Java bytecode~\cite{fink2012wala}.
Our goal was to obtain an upper bound on the number of redundant IFDS-result nodes that could be potentially removed by our analysis. The results are shown in Table~\ref{tab:dacapostat} in the Appendix.

@@

\subsection{An Example from the Scala Collections Library}\label{sec:collex}
The Scala collections library contains the trait \code{TraversableOnce} that is
shared by both collections and iterators over them. 
The \code{toArray} method of this trait
creates an array and copies the contents of the collection or iterator into it:
\inputMintedNoLNs{scala}{collectionsExample.scala}
%\otodo{Marianna, can you figure out how to get LaTeX to syntax-highlight this?} 
When \code{this} refers to an iterator rather than a collection, the call to \code{this.size}
extracts all elements of the iterator to count them.
 At the call to \code{copyToArray},
the iterator is already empty, so nothing is copied to the newly created array.
One could design an IFDS analysis to detect this kind of bug.

However, the implementation of \code{TraversableOnce.toArray} is actually correct because the above
code is guarded with a test: \code{if (this.isTraversableAgain) ...}
%\inputMintedNoLNs{scala}{isTraversable.scala}
When the \code{isTraversableAgain} method returns false, as it does for
an iterator, the \code{toArray} method uses a different
(less efficient) implementation. The bug report would therefore be a false positive. 
The \code{isTraversableAgain} method is easy to analyze: it returns the constant
true in a collection and the constant false in an iterator. However, in order
to eliminate the false positive bug report, an analysis would need to rule out infeasible
paths using correlated calls. Specifically, the following path triggers the bug, but
is infeasible: first, call \code{isTraversableAgain} on a collection, returning
true, then call \code{size} and \code{copyToArray} on an iterator. Our correlated
calls analysis could determine that this path is infeasible because it calls the
collection version of \code{isTraversableAgain} but the iterator versions of
\code{size} and \code{copyToArray}.

### 2015-03-08 sas2015/content/impl.tex
minor changes

\begin{restatable}[Soundness]{lemma}{sound}\label{lem:sound}
  Let $p=[\startmain,\,\dots,\,n]$ be a concrete execution path, and let $d\in D$.
  If $d\in M_F(p)(\varnothing),$
  then
  \begin{equation}
    d\in \backCC\left(\resultIDE(\transCC_R(P))\right)(n)\,.
  \end{equation}
\end{restatable}

@@

\begin{restatable}[Precision]{lemma}{precision}\label{lem:subsetifds}
    For an IFDS problem $P$ and all ${n\in N^*}$,
    \begin{equation}\label{eq:correct}
      \backCC\left(\resultIDE(\transCC_R(P))\right)(n)
      \subseteq
      \resultIFDS(P)(n)\,.
    \end{equation}
\end{restatable}

@@

\begin{restatable}{lemma}{onlycorrrec}\label{lem:onlycorrrec}
  Let $P$ be an IFDS problem. Then
  \begin{equation}
    \backCC\left(\resultIDE\left(\transCC_\rcc(P)\right)\right)=\backCC(\resultIDE\left(\transCC_R(P)\right))\,.
  \end{equation}
\end{restatable}

>>>>

\begin{restatable}[Soundness]{lemma}{sound}\label{lem:sound}
  Let $p=[\startmain,\,\dots,\,n]$ be a concrete execution path, and let $d\in D$.
  If $d\in M_F(p)(\varnothing),$
  then
  $
    d\in \backCC\left(\resultIDE(\transCC_R(P))\right)(n)\,.
  $
\end{restatable}

@@

\begin{restatable}[Precision]{lemma}{precision}\label{lem:subsetifds}
    For an IFDS problem $P$ and all ${n\in N^*}$,\\
    $
      \backCC\left(\resultIDE(\transCC_R(P))\right)(n)
      \subseteq
      \resultIFDS(P)(n)\,.
    $
\end{restatable}

@@

Let $P$ be an IFDS problemm.
\begin{restatable}{lemma}{onlycorrrec}\label{lem:onlycorrrec}
  $
    \backCC\left(\resultIDE\left(\transCC_\rcc(P)\right)\right)=\backCC(\resultIDE\left(\transCC_R(P)\right))\,.
  $
\end{restatable}

### 2015-03-08 sas2015/content/related.tex
minor changes

Sridharan et~al.'s improvements to the precision of pointer analysis
for Ja\-va\-Script programs~\cite{DBLP:conf/ecoop/SridharanDCST12} also
eliminate infeasible paths. However, instead of infeasible paths
between dynamically dispatched method calls, their approach eliminates
infeasible paths between reads and writes of different properties
(the JavaScript analogue of fields) of an object. Unlike in more static languages,
property names in JavaScript are first class values (strings). When a loop indexed by
property names updates the properties of objects, the loop body is extracted into a function
and analyzed context sensitively with a separate context for each property name.
The approach differs from ours in that it targets points-to analysis rather
than IFDS analyses, in that it targets infeasible paths due to different
property names rather than different dynamically dispatched methods,
and in that it employs context sensitivity to improve precision.

>>>>

Sridharan et~al.'s improvements to the precision of pointer analysis
for Ja\-va\-Script programs~\cite{DBLP:conf/ecoop/SridharanDCST12} also
eliminate infeasible paths. Instead of infeasible paths
between dynamically dispatched method calls, their approach eliminates
infeasible paths between reads and writes of different members
%(the JavaScript analogue of fields) 
of an object. 
%Unlike in more static languages,
%property names in JavaScript are first class values (strings). When a loop indexed by
%property names updates the properties of objects, the loop body is extracted into a function
%and analyzed context sensitively with a separate context for each property name.
The approach differs from ours in that it targets points-to analysis rather
than IFDS analyses, in that it targets infeasible paths due to different
property names rather than different dynamically dispatched methods,
and in that it employs context sensitivity to improve precision.
